Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-08-30 Manuscript/manuscript.Rmd
grammar edit

A growing number of studies have proposed computational tools used to identify the complete set of key molecular drivers of the formation of TAD boundaries. The classical approach is to build classification models that leverage the relationship between ChIP-seq genomic annotations and TAD boundary regions [@mourad2016computational; @huang2015predicting; @hong2017computational]. A key factor in these approaches becomes resolution. Resolution determines the bin sizes of the linear genome used to enumerate contacts caused by chromatin folding. Due to high sequencing cost, most available Hi-C datasets have relatively low resolution, in the kilobase range. This is in stark contrast to ChIP-seq experiments which annotate regions of the genome at hundreds of base pairs in resolution [??? REF, Bonev 2017?]. This resolution gap between 1D and 3D genomic maps, as well as a paucity of chromatin features below the level of TADs, has significantly limited our understanding of epigenomic regulation its potential link to chromatin architecture. However, simply increasing sequencing depth to obtain higher resolution could have unintended consequences since this could introduce sparsity in the number of contacts which TAD callers are often ill-equiped to handle [@cresswell2019spectraltad, @sexton2012three]. Increased resolution could also further exacerbate the class imbalance created from the proportionally much smaller number of TAD boundary regions compared to non-TAD boundary regions throughout the linear genome. Classifiers implemented on heavily imbalanced datasets are known to exhibit poor performances as the learned model is biased towards the majority class [@lusa2010class; @chen2013influence; @jeni2013facing]. It is unclear how simple data level solutions to class imbalance, such random under-sampling and random over-sampling affect TAD boundary prediction. Additionally, identifying the most informative feature engineering steps used to quantify the relationships between genomic anntotations and bins of varying resolutions remains an open question. Incorporating a distance measure has largely been ignored, but could offer a much more spatially descriptive feature space thereby improving predictive performance.

>>>>

A growing number of studies have proposed computational tools used to identify the complete set of key molecular drivers of the formation of TAD boundaries. The classical approach is to build classification models that leverage the relationship between ChIP-seq genomic annotations and TAD boundary regions [@mourad2016computational; @huang2015predicting; @hong2017computational]. A key factor in these approaches becomes resolution. Resolution determines the bin sizes of the linear genome used to enumerate contacts caused by chromatin folding. Due to high sequencing cost, most available Hi-C datasets have relatively low resolution, in the kilobase range. This is in stark contrast to ChIP-seq experiments which annotate regions of the genome at hundreds of base pairs in resolution [??? REF, Bonev 2017?]. This resolution gap between 1D and 3D genomic maps, as well as a paucity of chromatin features below the level of TADs, has significantly limited our understanding of epigenomic regulation its potential link to chromatin architecture. However, simply increasing sequencing depth to obtain higher resolution could have unintended consequences since this could introduce sparsity in the number of contacts which TAD callers are often ill-equiped to handle [@cresswell2019spectraltad, @sexton2012three]. Increased resolution could also further exacerbate the class imbalance created from the proportionally much smaller number of TAD boundary regions compared to non-TAD boundary regions throughout the linear genome. Classifiers implemented on heavily imbalanced datasets are known to exhibit poor performances as the learned model is biased towards the majority class [@lusa2010class; @chen2013influence; @jeni2013facing]. It is unclear how simple data level solutions to class imbalance, such random under-sampling and random over-sampling affect TAD boundary prediction. Additionally, identifying the most informative feature engineering steps used to quantify the relationships between genomic annotations and bins of varying resolutions remains an open question. Incorporating a distance measure has largely been ignored, but could offer a much more spatially descriptive feature space thereby improving predictive performance.
